---
title: "p8105_hw1_hl3978"
author: "Heather Lu"
date: "2025-09-18"
output: github_document
---

# Load libraries 

```{r}
library(tidyverse)
library(moderndive)
```

# Problem 1

## Load dataset

```{r}
data("early_january_weather")
```

## Dataset description

- The data set has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` Columns
- **Column names**: `r paste(names(early_january_weather))` 
- Importable variables include `time_hour` (timestamp), `temp` (F), `humid`, and `windspeed` (mph)
  - **Temperature range** (F) : `r paste(range(early_january_weather$temp), collapse = " to ")`
  - **Humidity range**: `r paste(range(early_january_weather$humid), collapse = " to ")`
  - **Windspeed range** (mph) : `r paste(range(early_january_weather$wind_speed), collapse = " to ")`

## temp vs. time_hour scatterplot

```{r}

#create scatterplot
temp_time_scatter <- ggplot(early_january_weather,
                            aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.8) +
  labs( x = "time (hourly)", y = "Temp (F)", color = "humidity",
        title = "Temperature over Time")

temp_time_scatter

# Export scatterplot in the project directory
ggsave("temp_time_scatter.png", plot = temp_time_scatter)

```

## Pattern description of temperature over time scatterplot

The plot shows that the temperature has a clear daily cycle - lower temperatures late night/early morning and higher temperatures midday/afternoon. There doesn't appear to be any clear patterns with humidity. In some cases, humidity seems to be associated with lower temperatures on certain days.

